---
title: "Lab 10 - Merging Data"
author: "Jackson Nahpi"
date: "November 2, 2017"
output: html_document
---

Using your own dataset (which may include more than one table) carry out the following data cleaning steps. Knit together the PDF document and commit both the Lab 10 RMD file and the PDF document to Git. Push the changes to GitHub so both documents are visible in your public GitHub repository. 

1. For your poster project, do you have multiple tables you'd like to join together to create your complete dataset? If so, describe what each table represents. 

I want to combine the tables for the sex vector and the overall dataset of all the other variables I want to include in my analysis.

2. What is/are your primary key(s)? If you have more than one table in your data, what is/are your foreign key(s)? Do your primary key(s) and foreign key(s) have the same name? If not, what does this mean for the way you need to specify potential data merges?

I don't have to match based on keys since my data each row represents a specific individual and is already organized with all of their traits present.

3. If you do not need to merge tables to create your final dataset, create a new dataset from your original dataset with a `grouped_by()` summary of your choice. You will use this separate dataset to complete the following exercises. 
```{r}
lab_dataset_v1 <- freshman_fall %>% 
  dplyr::select(parnt6, w1q8b, w1q8e, w1q8f, w1q8g, w1q8j, w1q8k, w1q10c)

lab_dataset_v2 <- freshman_fall %>% 
  dplyr::select(parnt13, w1q22a, w1q22e, w1q22f, w1q22g, w1q22m, w1q22n, w1q25c)

names(lab_dataset_v1) <- c(
          "prntsat6", 
          "checkhw6", 
          "rwrdfrgrds6", 
          "pnshfrgrds6", 
          "pnshdeviance6", 
          "artmusm6", 
          "scimusm6", 
          "educprogram6"
         )

names(lab_dataset_v2) <- c(
          "checkhw13", 
          "rwrdfrgrds13", 
          "pnshfrgrds13", 
          "pnshdeviance13", 
          "artmusm13", 
          "scimusm13", 
          "educprogram13",
          "prntsat13"
         )

lab_dataset_v1 <- transform(lab_dataset_v1,
                              prntsat6 = as.numeric(prntsat6), 
                              checkhw6 = as.numeric(checkhw6),
                              rwrdfrgrds6 = as.numeric(rwrdfrgrds6),
                              pnshfrgrds6 = as.numeric(pnshfrgrds6),
                              pnshdeviance6 = as.numeric(pnshdeviance6),
                              artmusm6 = as.numeric(artmusm6), 
                              scimusm6 = as.numeric(scimusm6),
                              educprogram6 = as.numeric(educprogram6)
                              )

lab_dataset_v2 <- transform(lab_dataset_v2,
                              prntsat13 = as.numeric(prntsat13),
                              checkhw13 = as.numeric(checkhw13),
                              rwrdfrgrds13 = as.numeric(rwrdfrgrds13),
                              pnshfrgrds13 = as.numeric(pnshfrgrds13),
                              pnshdeviance13 = as.numeric(pnshdeviance13),
                              artmusm13 = as.numeric(artmusm13),
                              scimusm13 = as.numeric(scimusm13),
                              educprogram13 = as.numeric(educprogram13)
                              )
```
I will be using these two altered datasets for this lab just to give practice in using the function, by no means do either of these datasets really show anything statistically significant. They are merely for my own learning.

If you are merging separate tables as part of your data manipulation process, are your keys of the same data type? If not, what are the differences? Figure out the appropriate coercion process(es) and carry out the steps below. 

4. Perform each version of the mutating joins (don't forget to specify the `by` argument) and print the results to the console. Describe what each join did to your datasets and what the resulting data table looks like. For those joining two separate datasets, did any of these joins result in your desired final dataset? Why or why not?
```{r}
print(left_join(x = lab_dataset_v1, y = lab_dataset_v2, by = c("prntsat6" = "prntsat13")))

print(right_join(x = lab_dataset_v1, y = lab_dataset_v2, by = c("prntsat6" = "prntsat13")))
```

For the first left join the two tables were joined depending on whether both prntsat6 had the same values as prntsat13 giving us all the columns in order of prntsat6 ordering. The second join, the right_join, does the same thing as the left join but using the right column (prntsat13) as the lead column of the join.

5. Do the same thing with the filtering joins. What was the result? Give an example of a case in which a `semi_join()` or an `anti_join()` might be used with your primary dataset

```{r}
semi_join(x = lab_dataset_v1, y = lab_dataset_v2, by = c("prntsat6" = "prntsat13"))
anti_join(x = lab_dataset_v1, y = lab_dataset_v2, by = c("prntsat6" = "prntsat13"))
```
6. What happens when you apply the set operations joins to your tables? Are these functions useful for you for this project? Explain why or why not. If not, give an example in which one of them might be usefully applied to your data. 

These joins are most likely not useful to me in this project since each rows variables are all related to a singular person therefore it isn't very useful to manipulate each columns row values individually using join functions.


7. If you have any reason to compare tables, apply `setequal()` below. What were the results? 
```{r}
dplyr::setequal(x = lab_dataset_v1, y = lab_dataset_v2)
```

8. What is the purpose of binding data and why might you need to take extra precaution when carrying out this specific form of data merging? If your data requires any binding, carry out the steps below and describe what was accomplished by your merge.

Binds are used to combine two data frames together and depending on which way you bind you columns(or rows) you can either be adding the left or the right of the data frame. A precaution when conducting this type of merge is to make sure that the rows or columns are in the same order in each of the data frames. What I mean by this is that by adding one column to another column the rows may be linked to specific observations an by blindly merging them you may accidentally mess up the variables related to each observation. My data does require merging using colbind but I will do this below on question 10.

9. Do you need to merge multiple tables together using the same type of merge? If so, utilize the `reduce()` function from the `purrr` package to carry out the appropriate merge below. 

No I do not require the merging of multiple tables

10. Are there any other steps you need to carry out to further clean, transform, or merge your data into one, final, tidy dataset? If so, describe what they are and carry them out below.
```{r}
freshman_dataset <- freshman_fall %>% 
  dplyr::select(caseid, college, parnt6, parnt13, parntlyr, w1q8b, w1q8e, w1q8f, w1q8g, w1q8j, w1q8k, w1q10c, w1q22a, w1q22e, w1q22f, w1q22g, w1q22m, w1q22n, w1q25c, w1q38a, w1q38e, w1q38f, w1q38g, w1q38m, w1q87b, w1q90, w1q91a, w1q91b, w1q91c, w1q91d, w1q91e)

sex_vector <- ifelse(freshman_fall$sex == "M", 0, 1)

sex_vector <- data.frame(sex_vector)

freshman_dataset <- bind_cols(sex_vector, freshman_dataset)

freshman_dataset <- na.omit(freshman_dataset)

names(freshman_dataset) <- c("sex",
          "caseid",
          "college",
          "prntsat6", 
          "prntsat13", 
          "prntsatlyr", 
          "checkhw6", 
          "rwrdfrgrds6", 
          "pnshfrgrds6", 
          "pnshdeviance6", 
          "artmusm6", 
          "scimusm6", 
          "educprogram6",
          "checkhw13", 
          "rwrdfrgrds13", 
          "pnshfrgrds13", 
          "pnshdeviance13", 
          "artmusm13", 
          "scimusm13", 
          "educprogram13",
          "checkhwlyr", 
          "rwrdfrgrdslyr", 
          "pnshfrgrdslyr", 
          "pnshdeviancelyr", 
          "musmlyr", 
          "educprogramlyr", 
          "currntcolaspir", 
          "probfincol1yr",
          "probfincol2yr",
          "probfincol",
          "probextraeduc",
          "probfingraddeg"
         )

freshman_dataset <- transform(freshman_dataset,
                              caseid = as.numeric(caseid),
                              prntsat6 = as.numeric(prntsat6),
                              prntsat13 = as.numeric(prntsat13),
                              prntsatlyr = as.numeric(prntsatlyr)
                              )

freshman_dataset <- transform(freshman_dataset,
                              checkhw6 = as.numeric(checkhw6),
                              rwrdfrgrds6 = as.numeric(rwrdfrgrds6),
                              pnshfrgrds6 = as.numeric(pnshfrgrds6),
                              pnshdeviance6 = as.numeric(pnshdeviance6),
                              artmusm6 = as.numeric(artmusm6), 
                              scimusm6 = as.numeric(scimusm6),
                              educprogram6 = as.numeric(educprogram6)
                              )

freshman_dataset <- transform(freshman_dataset,
                              checkhw13 = as.numeric(checkhw13),
                              rwrdfrgrds13 = as.numeric(rwrdfrgrds13),
                              pnshfrgrds13 = as.numeric(pnshfrgrds13),
                              pnshdeviance13 = as.numeric(pnshdeviance13),
                              artmusm13 = as.numeric(artmusm13),
                              scimusm13 = as.numeric(scimusm13),
                              educprogram13 = as.numeric(educprogram13)
                              )
                              
freshman_dataset <- transform(freshman_dataset,
                              checkhwlyr = as.numeric(checkhwlyr),
                              rwrdfrgrdslyr = as.numeric(rwrdfrgrdslyr),
                              pnshfrgrdslyr = as.numeric(pnshfrgrdslyr),
                              pnshdeviancelyr = as.numeric(pnshdeviancelyr),
                              musmlyr = as.numeric(musmlyr),
                              educprogramlyr = as.numeric(educprogramlyr)
                              )

freshman_dataset <- transform(freshman_dataset,
                              currntcolaspir = as.numeric(currntcolaspir),
                              probfincol1yr = as.numeric(probfincol1yr),
                              probfincol2yr = as.numeric(probfincol2yr),
                              probfincol = as.numeric(probfincol),
                              probextraeduc = as.numeric(probextraeduc),
                              probfingraddeg = as.numeric(probfingraddeg)
                              )


freshman_dataset <- filter(freshman_dataset, 
                           checkhw6 < 6 & 
                             rwrdfrgrds6 < 6 &
                             pnshfrgrds6 < 6 &
                             pnshdeviance6 < 6 &
                             artmusm6 < 6 &
                             scimusm6 < 6 &
                             educprogram6 < 6 &
                             checkhw13 < 6 &
                             rwrdfrgrds13 < 6 &
                             pnshfrgrds13 < 6 &
                             pnshdeviance13 < 6 & 
                             artmusm13 < 6 &  
                             scimusm13 < 6 &  
                             educprogram13 < 6 &
                             checkhwlyr < 6 & 
                             rwrdfrgrdslyr < 6 & 
                             pnshfrgrdslyr < 6 &  
                             pnshdeviancelyr < 6 & 
                             musmlyr < 6 & 
                             probfincol1yr < 97 &
                             probfincol2yr < 97 & 
                             probfincol < 97 & 
                             probextraeduc < 97 & 
                             probfingraddeg < 97
                             )

freshman_dataset <- freshman_dataset %>%
  filter(educprogramlyr != 97 & educprogramlyr != 98) %>%
  mutate(educprogramlyr = ifelse(test = educprogramlyr >= 1, 
                             yes = 1, 
                             no = 0))


freshman_dataset <- mutate(freshman_dataset,
                           musm6 = round((freshman_dataset$artmusm6 + 
                                            freshman_dataset$scimusm6)/2, 0))
  
freshman_dataset <- mutate(freshman_dataset, 
                           musm13 = round((freshman_dataset$artmusm13 + 
                                             freshman_dataset$scimusm13)/2, 0))

freshman_dataset$artmusm6 <- NULL
freshman_dataset$scimusm6 <- NULL
freshman_dataset$artmusm13 <- NULL
freshman_dataset$scimusm13 <- NULL
```

I first started out by binding my overall table to the sex vector. I then went through removing all NAs and conversion of all variables into numeric form by using the 'transform' function. Thirdly, I went through and filtered out all variable values that were related to no response or don't knows. Fourthly, I mutated my dataset by combing the artmusm and scimusm variables for ages 6 and 13 into 2 variables (reduced 4 variables down to 2) and then devided those new variables musm6(13) by 2, this way I can keep the original measureing rubric that was used in the original study. Lastly, I removed the 4 variables that were reduced to 2 variables since they were no longer needed.

