---
title: "variables_list"
author: "Jackson Nahpi"
date: "October 25, 2017"
output: html_document
---
#packages
```{R}
library(foreign)
library(haven)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(memisc)
library(purrr)
library(ggthemes)
library(knitr)
library(gmodels)
library(GGally)
library(broom)
library(vcd)
library(ggtern)
library(stargazer)
```


```
Change names on the independent variables.
Second line removes title
Third line changes the column names
Fourth, alters and seperates columns
```{r}
covariate.labels = c("","","")

dep.var.labels.include = FALSE

column.labels = c("Good", "Better")

column.separate = c(2, 2)
```


Code to Generate the Model Table in Latex format use quick latex to render
```{r, echo = FALSE, warning = FALSE}
stargazer(model_8, 
          model_9, 
          model_10, 
          type = 'text', 
          style = "qje", 
          title = "What is Significant To Optimism in Graduate Degree Attainment", 
          digits = 3, 
          out = "table.txt", 
          summary = TRUE, 
          single.row = TRUE,
          covariate.labels = c("Rewarded For Grades(2)",
                               "(3)",
                               "(4)",
                               "(5)",
                               "Homework Gets Checked(2)",
                               "(3)",
                               "(4)",
                               "(5)",
                               "Disobedience Punishment(2)",
                               "(3)",
                               "(4)",
                               "(5)",
                               "Sex",
                               "Ethnicity(2)",
                               "(3)",
                               "(4)",
                               "Museum Visits(2)",
                               "(3)",
                               "(4)",
                               "(5)",
                               "College Choice(2)",
                               "(3)",
                               "Chance Of Attaining Extra Education",
                               "Chance Of Finishing College"),
          column.labels = c("Model 1",
                            "Model 2",
                            "Model 3"),
          dep.var.labels = c("Chance of Attaining a Graduate Degree"))
```

```{r mylatextable, results = "asis"}
stargazer(model_10, type = 'latex')

```


```{r}
freshman_dataset4$ethnicity <- mapvalues(freshman_dataset_4$ethnicity,
c('1', '2'),
c("black","white"))
```

```{r,echo = FALSE, warning = FALSE}
mosaic(data = freshman_dataset_v4,
       ~ sex + college+ ethnicity,
       shade = TRUE, 
       legend = TRUE,
       main = "Effects of Race and Gender on College Decisions", 
       zero_split = TRUE, 
       spacing_args = list(start = 0.5, 
                           rate = 1.5),
       labeling= labeling_border(rot_labels = c(0,
                                                0,
                                                0,
                                                0),
                                 rot_varnames = c(0,
                                                  0,
                                                  0,
                                                  0),
                                 just_labels = c("center", 
                                                 "left", 
                                                 "left", 
                                                 "right"), 
                                 set_varnames = list(sex = "Gender", 
                                                     ethnicity = "Race",
                                                     college = "College Choice"),
                                 set_labels = lnames,
                                 offset_varnames = c(0.5,
                                                     3,
                                                     0,
                                                     2.25)),
       spacing = spacing_dimequal(c(0.5,
                                    0.15,
                                    0.15)),
       margins = c(left = 7, 
                   top = 3.5, 
                   right = 7), 
       gp_args = list(interpolate = 1:4))
```


#Possible Addons
```{r}
direction = c("v","h","v")
#Adjusts the location of where different variables are placed in the mosaic plot
sub = "Male"
#Adds a title wherever the location specified

split_vertical = c(TRUE, FALSE, TRUE)
#alters how the mosaic is shown by rotating the matrix


mosaic(~ college + ethnicity | sex, data = mosaic_subset, split_vertical = c(TRUE, TRUE, FALSE), legend = TRUE)
#WHY
```



```{r}
mosaic_table <- table(mosaic_subset)
mosaicplot(table(mosaic_table), color = 1:3, las = 1, main = "test", xlab="whatevs", ylab="else")
mtext(side = 1, "Category1", line = 0.5, col="green")
mtext(side = 1, "Category1", line = 1, col="blue")
mtext(side = 1, "Category1", line = 2, col="red")
mtext(side = 2, "Category2", line = -1, col="purple")
```

#How to knit code
```{r}
rmarkdown::render('Lab 8 - Data cleaning.rmd', output_format = 'pdf_document')
knitr::opts_chunk$set(echo = TRUE)
```

#How to conduct multiple regression with interaction terms
```{r}
model_8 <- lm(probfingraddeg ~ factor(rwrdfrgrdslyr) + 
                factor(checkhwlyr) + 
                factor(pnshdeviancelyr) +
                factor(sex) +
                factor(ethnicity) +
                factor(musmlyr) +
                factor(college) +
              factor(sex):probfincol,
              data = freshman_dataset_v4)
summary(model_8)
```

#data sets
```{R}
freshman_fall<-read.dta("nlsfw1p.dta/nlsfw1p.dta")
freshman_spring<-read.dta("nlsfw2p.dta/nlsfw2p.dta")
sophomore_spring<-read.dta("nlsfw3p.dta/nlsfw3p.dta")
junior<-read.dta("nlsfw4p.dta/nlsfw4p.dta")
senior<-read.dta("nlsfw5p.dta/nlsfw5p.dta")
```

Experimental
```{r}
test_freshman_dataset <- freshman_fall %>% 
  dplyr::select(caseid, college, parnt6, parnt13, parntlyr, w1q8b, w1q8e, w1q8f, w1q8g, w1q8j, w1q8k, w1q10c, w1q22a, w1q22e, w1q22f, w1q22g, w1q22m, w1q22n, w1q25c, w1q38a, w1q38e, w1q38f, w1q38g, w1q38m, w1q87b, w1q90, w1q91a, w1q91b, w1q91c, w1q91d, w1q91e)

test_sex_vector <- ifelse(freshman_fall$sex == "M", 0, 1)

test_sex_vector <- data.frame(test_sex_vector)

race_vector <- dplyr::recode(freshman_fall$w1qzeth, "B"=1,"W"=2,"A"=3,"H"=4)

race_vector <- data.frame(race_vector)

sex_and_ethnicity_vector <- bind_cols(test_sex_vector, race_vector)

test_freshman_dataset <- bind_cols(sex_and_ethnicity_vector, test_freshman_dataset)

test_freshman_dataset <- na.omit(test_freshman_dataset)

names(test_freshman_dataset) <- c("sex",
          "ethnicity",
          "caseid",
          "college",
          "prntsat6", 
          "prntsat13", 
          "prntsatlyr", 
          "checkhw6", 
          "rwrdfrgrds6", 
          "pnshfrgrds6", 
          "pnshdeviance6", 
          "artmusm6", 
          "scimusm6", 
          "educprogram6",
          "checkhw13", 
          "rwrdfrgrds13", 
          "pnshfrgrds13", 
          "pnshdeviance13", 
          "artmusm13", 
          "scimusm13", 
          "educprogram13",
          "checkhwlyr", 
          "rwrdfrgrdslyr", 
          "pnshfrgrdslyr", 
          "pnshdeviancelyr", 
          "musmlyr", 
          "educprogramlyr", 
          "currntcolaspir", 
          "probfincol1yr",
          "probfincol2yr",
          "probfincol",
          "probextraeduc",
          "probfingraddeg"
         )

test_freshman_dataset <- test_freshman_dataset %>%
  filter(educprogramlyr != 97 & educprogramlyr != 98) %>%
  mutate(educprogramlyr = ifelse(test = educprogramlyr >= 1, 
                             yes = 1, 
                             no = 0))

test_freshman_dataset <- transform(test_freshman_dataset,
                              caseid = as.numeric(caseid),
                              prntsat6 = as.numeric(prntsat6),
                              prntsat13 = as.numeric(prntsat13),
                              prntsatlyr = as.numeric(prntsatlyr)
                              )

test_freshman_dataset <- transform(test_freshman_dataset,
                              checkhw6 = as.numeric(checkhw6),
                              rwrdfrgrds6 = as.numeric(rwrdfrgrds6),
                              pnshfrgrds6 = as.numeric(pnshfrgrds6),
                              pnshdeviance6 = as.numeric(pnshdeviance6),
                              artmusm6 = as.numeric(artmusm6), 
                              scimusm6 = as.numeric(scimusm6),
                              educprogram6 = as.numeric(educprogram6)
                              )

test_freshman_dataset <- transform(test_freshman_dataset,
                              checkhw13 = as.numeric(checkhw13),
                              rwrdfrgrds13 = as.numeric(rwrdfrgrds13),
                              pnshfrgrds13 = as.numeric(pnshfrgrds13),
                              pnshdeviance13 = as.numeric(pnshdeviance13),
                              artmusm13 = as.numeric(artmusm13),
                              scimusm13 = as.numeric(scimusm13),
                              educprogram13 = as.numeric(educprogram13)
                              )
                              
test_freshman_dataset <- transform(test_freshman_dataset,
                              checkhwlyr = as.numeric(checkhwlyr),
                              rwrdfrgrdslyr = as.numeric(rwrdfrgrdslyr),
                              pnshfrgrdslyr = as.numeric(pnshfrgrdslyr),
                              pnshdeviancelyr = as.numeric(pnshdeviancelyr),
                              musmlyr = as.numeric(musmlyr),
                              educprogramlyr = as.numeric(educprogramlyr)
                              )

test_freshman_dataset <- transform(test_freshman_dataset,
                              currntcolaspir = as.numeric(currntcolaspir),
                              probfincol1yr = as.numeric(probfincol1yr),
                              probfincol2yr = as.numeric(probfincol2yr),
                              probfincol = as.numeric(probfincol),
                              probextraeduc = as.numeric(probextraeduc),
                              probfingraddeg = as.numeric(probfingraddeg)
                              )


test_freshman_dataset <- filter(test_freshman_dataset, 
                           checkhw6 < 6 & 
                             rwrdfrgrds6 < 6 &
                             pnshfrgrds6 < 6 &
                             pnshdeviance6 < 6 &
                             artmusm6 < 6 &
                             scimusm6 < 6 &
                             educprogram6 < 6 &
                             checkhw13 < 6 &
                             rwrdfrgrds13 < 6 &
                             pnshfrgrds13 < 6 &
                             pnshdeviance13 < 6 & 
                             artmusm13 < 6 &  
                             scimusm13 < 6 &  
                             educprogram13 < 6 &
                             checkhwlyr < 6 & 
                             rwrdfrgrdslyr < 6 & 
                             pnshfrgrdslyr < 6 &  
                             pnshdeviancelyr < 6 & 
                             musmlyr < 6 & 
                             probfincol1yr < 97 &
                             probfincol2yr < 97 & 
                             probfincol < 97 & 
                             probextraeduc < 97 & 
                             probfingraddeg < 97
                             )

test_freshman_dataset <- mutate(test_freshman_dataset,
                           musm6 = round((test_freshman_dataset$artmusm6 + 
                                            test_freshman_dataset$scimusm6)/2, 0))
  
test_freshman_dataset <- mutate(test_freshman_dataset, 
                           musm13 = round((test_freshman_dataset$artmusm13 + 
                                             test_freshman_dataset$scimusm13)/2, 0))

test_freshman_dataset$artmusm6 <- NULL
test_freshman_dataset$scimusm6 <- NULL
test_freshman_dataset$artmusm13 <- NULL
test_freshman_dataset$scimusm13 <- NULL
```

